---
title: "W7_CW_notes"
author: "Jasmine Chang"
format: html
editor: visual
---

# Week 7 Lecture 1 Notes

## Maps Part 1

Various maps packages:

{maps} Simple polygon basemaps (country, state, county) No {sf} Full-featured spatial data (GIS-grade) No {rnaturalearth} World/country/state polygons as sf objects No {maptiles} Download satellite/street tile basemaps No {leaflet} Interactive HTML maps No {tmap} Thematic maps, static and interactive No {ggmap}

### Load Libraries

```{r}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
```

### Read in Data

```{r}
# Population in California by county
popdata <- read_csv(here("Week_7", "data", "CApopdata.csv"))

# Seastar counts at field sites in California
stars <- read_csv(here("Week_7", "data", "stars.csv"))
```

###The **map_data()** package

map_data()-World

```{r}
world <- map_data("world")

head(world)
```

map_data()-USA

```{r}
usa <- map_data("usa")

head(usa)
```

map_data() — A Country

```{r}
italy <- map_data("italy")

head(italy)
```

map_data() — US States

```{r}
states <- map_data("state")

head(states)
```

map_data() — US Counties

```{r}
counties <- map_data("county")

head(counties)
```

###Structure of the Data \* long — longitude. West of the prime meridian is negative. \* lat — latitude. \* order — the order in which ggplot connects the dots (polygon vertices). \* region / subregion — the region or subregion a set of points surrounds. \* group — This one is critical! Controls whether adjacent points are connected by lines. Points in the same group get connected; different groups cause ggplot to “lift the pen.” **warning**: Forgetting group = group in your aesthetic is the #1 cause of “crazy line” plots. You will do this. Everyone does.

####Make a Map of the World

```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group))
```

What Happens Without group = group?

```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat))
# essentially tries to connect all points to eachother
```

Add Color to Lines and Fill

```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group),
               color = "red")
```

Fill by Region

```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group,
                   fill = region),
               color = "black") +
  guides(fill = "none") #get rid of legend, because that would be too much
```

Change the theme

```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group,
                   fill = region),
               color = "black") +
  guides(fill = "none") +
  theme_minimal()
```

Make the Ocean Blue

```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group,
                   fill = region),
               color = "black") +
  theme_minimal() +
  guides(fill = "none") +
  theme(panel.background =
          element_rect(fill = "lightblue"))
```

Change the Map Projection The earth is not flat, but we visualize it in 2D — there are many projections to choose from. See {mapproj} for available options. Use a Mercator projection:

```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group,
                   fill = region),
               color = "black") +
  theme_minimal() +
  guides(fill = "none") +
  theme(panel.background =
          element_rect(fill = "lightblue")) +
  coord_map(projection = "mercator", # there are different projections to choose from
            xlim = c(-180, 180))
```
Use a sinusoidal projection:

```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group,
                   fill = region),
               color = "black") +
  theme_minimal() +
  guides(fill = "none") +
  theme(panel.background =
          element_rect(fill = "lightblue")) +
  coord_map(projection = "sinusoidal",
            xlim = c(-180, 180))
```

Map of Just California
Filter the states dataset to California:
```{r}
CA_data <- states |>
  filter(region == "california")
```

Make the map:
```{r}
ggplot() +
  geom_polygon(data = CA_data,
               aes(x = long, y = lat,
                   group = group),
               fill = "gold")+
  labs(title = "State of California",
       x="Longitude",
       y="Latitude")+
  theme(panel.background =
          element_rect(fill = "lightblue"))+
  coord_map(projection = "mercator")+ # makes it more to scale
  theme_classic()

```

Adding Multiple Layers of Data
Plot the population of every county in California.
First, we need county-level polygons and join them to our population dataset.

```{r}
head(counties)[1:3, ]

```

```{r}
head(popdata)
```
Wrangle the Data
Make column names match before joining.

```{r}
CApop_county <- popdata |>
  select("subregion" = County, Population) |>   # rename to match counties
  inner_join(counties) |>
  filter(region == "california")                 # some county names repeat in other states

head(CApop_county)
```

Map of CA Population by County

```{r}
ggplot() +
  geom_polygon(data = CApop_county,
               aes(x = long, y = lat,
                   group = group,
                   fill = Population),
               color = "black") +
  coord_map() +
  theme_void() # pretty, but hard to see differences in populations
```
Log Scale for Easier Interpretation
```{r}
ggplot() +
  geom_polygon(data = CApop_county,
               aes(x = long, y = lat,
                   group = group,
                   fill = Population),
               color = "black") +
  coord_map() +
  theme_void() +
  scale_fill_gradient(transform = "log10") 
#transform = “log10” compresses the color scale so that less-populated counties are still visible alongside Los Angeles.
```
### Stars!!

Add a Layer of Points
The stars dataset has seastar counts (per m²) at field sites in California.

```{r}
head(stars)
```

Overlay Seastar Sites

```{r}
ggplot() +
  geom_polygon(data = CApop_county,
               aes(x = long, y = lat,
                   group = group,
                   fill = Population),
               color = "black") +
  geom_point(data = stars,
             aes(x = long,
                 y = lat)) +
  coord_map() +
  theme_void() +
  scale_fill_gradient(transform = "log10")
```
Size Points by Seastar Count
```{r}
ggplot() +
  geom_polygon(data = CApop_county,
               aes(x = long, y = lat,
                   group = group,
                   fill = Population),
               color = "black") +
  geom_point(data = stars,
             aes(x = long, y = lat,
                 size = star_no)) +
  coord_map() +
  theme_void() +
  scale_fill_gradient(transform = "log10")+
    labs(size = "# Stars/m²") #better legend labels

ggsave(here("Week_7", "output", "CApop.pdf")) #saves most recent plot by default
```
## {sf}

Why {sf}?
The {sf} (“simple features”) package is the current standard for spatial data in R.
* Spatial objects are just data frames — all tidyverse verbs work on them
* Handles projections, CRS transformations, spatial joins natively
* geom_sf() replaces geom_polygon() — no need to manage group manually
* Compatible with a rich ecosystem: {tigris}, {rnaturalearth}, {maptiles}, {leaflet}

Load Libraries

```{r}
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

```
{rnaturalearth} — World Map with geom_sf()
can make maps with a lot less code
```{r}
world_sf <- ne_countries(scale = "medium",
                         returnclass = "sf")

ggplot(world_sf) +
  geom_sf(fill = "lightgray",
          color = "white") +
  theme_void()
```
Projections with coord_sf()
coord_sf() handles projections using standard EPSG codes — much more rigorous than {mapproj}:
Common CRS choices: "+proj=robin" (Robinson), 4326 (WGS84 / GPS), 3857 (Web Mercator).
Learn more at map chapter in ggplot2 book.
```{r}
ggplot(world_sf) +
  geom_sf(fill = "lightgray", color = "white") +
  coord_sf(crs = "+proj=robin") + # Robinson projection
  theme_void() +
  labs(title = "Robinson Projection")
```
Total Awesome R Package — {ggdogs} 🐾

```{r}
library(ggdogs)
ggplot(mtcars) +
  geom_dog(aes(mpg, wt), dog = "pug", size = 8)
```

