---
title: "First Quarto Doc"
author: "Jasmine Chang"
format: html
editor: visual
---

### My document (notes)

```{r}
#| warning: false    # suppress warnings in all chunks
#| message: false   # suppress package loading messages
#| echo: false      # hide code by default (show results only)
#| eval: true       # run all chunks (the default)
1+1
```

# In your prose (not a code chunk), write:

# backtick + r + space + expression + backtick

# Examples:

# The dataset has `nrow(penguins)` rows.

# The mean is `round(mean(x, na.rm = TRUE), 1)`.

# Analysis run on `Sys.Date()`.

### Activity: Inline R code

```{r}
## load libraries ##
library(palmerpenguins)
library(here)
library(tidyverse)
library(rsconnect)
## load data ##
glimpse(penguins)
## publish code ##
rsconnect::deployDoc(
  doc      = here("Week_6","scripts", "my_analysis.qmd"),  # path to your .qmd
  appTitle = "OCN 682 - Your Name - Week 6"        # give it a clear title
)
```

There are **3** total species of penguins. The maximum bill length is `r max(penguins$bill_length_mm, na.rm=TRUE)` and the minimum bill length is `r min(penguins$bill_lengh_mm, na.rm=TRUE)`. The mean bill length is `round(mean(penguins$bill_length_mm, na.rm=TRUE),1)`.
